---
title: "project1_crm"
output:
  html_document:
    toc: true
    toc_depth: 2
    keep_tex: true
---

# Set Working Directory
```{r Set Working Directory}
setwd("C:/Users/camil/OneDrive - Baylor College of Medicine/Desktop/R Work/R Directory/BCM Work")
```

# Load in Required Packages
```{r Load in Required Packages, warning=FALSE, message=FALSE}
library(Seurat)
library(dplyr)
```

# Load in Data
```{r Load in Data}
data <- Read10X(data.dir = "C:/Users/camil/OneDrive - Baylor College of Medicine/Desktop/R Work/R Directory/BCM Work/5k_mouse_brain_CNIK_3pv3_filtered_feature_bc_matrix/filtered_feature_bc_matrix")
```

# Make a Seurat Object
```{r Make a Seurat Object}
# We set the minimum number of features to 50 since the neurons have a lower RNA abundance than glial cells
data_seurat <- CreateSeuratObject(counts = data, 
                           project = "pbmc3k", 
                           min.cells = 3, 
                           min.features = 50)
```

# Make Column in Metadata for Mitochondrial Percentage
```{r Mitochondrial Percentage}
data_seurat[["percent.mt"]] <- PercentageFeatureSet(data_seurat, pattern = "^mt-")
```

# Violin Plots of Features, Counts, and Mitochondrial Percentage
```{r Violin Plots, warning=FALSE}
VlnPlot(data_seurat, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size=.2)
# Can see that the violin plot for number of features has a bimodal distribution
# This is due to the differences in RNA abundance between neurons and glial cells
```

# Make a Vector for Number of Features from Metadata
```{r nFeature Vector }
a = data_seurat[["nFeature_RNA"]]$nFeature_RNA
```

# Find 99th Percentile for Number of Features
```{r 99th Percentile}
cat(quantile(a, 0.99))
```

# Made a Subset of Seurat Object
```{r QC Subset of Seurat Object}
data_seurat <- subset(data_seurat, subset = nFeature_RNA > 50 & nFeature_RNA < 5139 & percent.mt < 5)
# We want to have less than 5% mitochondria to exclude dead cells
# We want cells with enough detected genes, so more than 50 and less than the 99% quantile we got earlier
# We exclude above 99% quantile so we can exclude potential doublets

dim(data_seurat)
# This shows there are 22738 genes present and 5716 cells passed the quality control filter we set
```

# Normalized the Subset
```{r Log Normalize Subset, message=FALSE}
data_seurat <- NormalizeData(data_seurat, normalization.method = "LogNormalize", scale.factor = 10000)
```

# Identify Variable Genes
```{r Identify Feature Outliers, message=FALSE}
data_seurat <- FindVariableFeatures(data_seurat, selection.method = "vst", nfeatures = 2000)
# chose top 2000 most variable genes using a mean variability plot
```

# Make a Vector of Names of Variable Genes 
```{r Vector of Variable Genes}
all.genes <- rownames(data_seurat)
# This vector should name the top 2000 variable genes identified
```

# Scale and Center Expression Values of Genes
```{r Scale and Center Gene Expression, message=FALSE}
data_seurat <- ScaleData(data_seurat, features = all.genes)
```

# PCA
```{r PCA}
# use VariableFeatures to select only the highly variable genes for PCA
data_seurat <- RunPCA(data_seurat, features = VariableFeatures(object = data_seurat))

# Plot PCA results
DimPlot(data_seurat, reduction = "pca")
```

# Made an Elbow Plot
```{r Elbow Plot}
# Assess percentage standard deviation explained by different PCs
ElbowPlot(data_seurat, ndims=50)
```

# JackStraw Analysis and Plot
```{r JackStraw, warning=FALSE}
# Pick number of PCs for a permutation-based test 
data_seurat <- JackStraw(data_seurat, num.replicate = 20)
# Did permutation test 20 times on PCs
data_seurat <- ScoreJackStraw(data_seurat, dims = 1:20)
# Assigns p-values to the PCs to identify PCs that actually explain the variance
JackStrawPlot(data_seurat, dims = 1:20, ymax = .35)

# QQplot for significance of each PC
# Set y-axis to .35 so you can see separation of PCs in the JackStrawPlot
# Dots above diagonal line show PCs that explain variance
# The further away the PC curve is from standard distribution, the more that PC explains true variance
```

# Clustering and Visualization of tSNE/UMAP
```{r Form Clusters, message=FALSE}
# find nearest neighbors using the first 10 PCs
data_seurat <- FindNeighbors(data_seurat, dims = 1:10) 

# find clusters of cells
data_seurat <- FindClusters(data_seurat, resolution = 0.5, verbose=FALSE)
# set resolution at .5 to find "larger" clusters
```

# UMAP
```{r UMAP, warning=FALSE, message=FALSE}
# dims sets the number of PCs used for finding neighbors/clusters
data_seurat_5 <- RunUMAP(data_seurat, dims = 1:5)
# we think 10 PCs would give the best results for UMAP
data_seurat <- RunUMAP(data_seurat, dims = 1:10)
data_seurat_15 <- RunUMAP(data_seurat, dims = 1:15)
data_seurat_20 <- RunUMAP(data_seurat, dims = 1:20)

# Plot UMAP results with different number of PCs
DimPlot(data_seurat_5, reduction = "umap", label=T)
DimPlot(data_seurat, reduction = "umap", label=T)
DimPlot(data_seurat_15, reduction = "umap", label=T)
DimPlot(data_seurat_20, reduction = "umap", label=T)
```

# tSNE with Multiple Perplexities
```{r tSNE}
# perplexity controls size of neighborhood around each data point, should be between 5 and 50
# perplexity should be n/100, there are 5716 cells so ideal perplexity is probably about 50
# changed "perplecity" to "perplexity" in code

data_seurat_5 <- RunTSNE(data_seurat, dims = 1:10, perplexity = 5)
data_seurat_10 <- RunTSNE(data_seurat, dims = 1:10, perplexity = 10)
data_seurat_20 <- RunTSNE(data_seurat, dims = 1:10, perplexity = 20)
data_seurat_30 <- RunTSNE(data_seurat, dims = 1:10, perplexity = 30)
data_seurat_40 <- RunTSNE(data_seurat, dims = 1:10, perplexity = 40)
data_seurat_50 <- RunTSNE(data_seurat, dims = 1:10, perplexity = 50)

DimPlot(data_seurat_5, reduction = "tsne",label=T)
DimPlot(data_seurat_10, reduction = "tsne",label=T)
DimPlot(data_seurat_20, reduction = "tsne",label=T)
DimPlot(data_seurat_30, reduction = "tsne",label=T)
DimPlot(data_seurat_40, reduction = "tsne",label=T)
DimPlot(data_seurat_50, reduction = "tsne",label=T)
```

# Identify Marker Genes
```{r Identify marker genes, message=FALSE}
cluster1.markers <- FindMarkers(data_seurat, ident.1 = 1, min.pct = 0.25, verbose=FALSE)

# see genes in each cluster
head(cluster1.markers)

# view  differentially marked genes in each class of the dataset 
all.markers <- FindAllMarkers(data_seurat, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 1, verbose=FALSE)

table_markers <- all.markers[, c("gene", "p_val", "cluster", "avg_log2FC", "pct.1", "pct.2", "p_val_adj")]

# want the two genes with highest log2FC for each cluster
table_markers <- table_markers %>%
  group_by(cluster) %>%
  slice_max(n=2, order_by = avg_log2FC)

# top genes of each cluster
knitr::kable(table_markers)  


# list of top genes
table_markers$gene
```


# Heatmap of Genes by Cluster
```{r Heatmap, out.height="125%"}
DoHeatmap(object = data_seurat, features=table_markers$gene, group.by="seurat_clusters", size=3)
```

# UMAP for Top Genes of Each Cluster
```{r UMAP of Top Genes}
FeaturePlot(data_seurat,reduction="umap",features=table_markers$gene[1:2],order=F,cols=c("#d9d9d933","#bd0026"), ncol = 2, label=TRUE)
FeaturePlot(data_seurat,reduction="umap",features=table_markers$gene[3:4],order=F,cols=c("#d9d9d933","#bd0026"), ncol = 2, label=TRUE)
FeaturePlot(data_seurat,reduction="umap",features=table_markers$gene[5:6],order=F,cols=c("#d9d9d933","#bd0026"), ncol = 2, label=TRUE)
FeaturePlot(data_seurat,reduction="umap",features=table_markers$gene[7:8],order=F,cols=c("#d9d9d933","#bd0026"), ncol = 2, label=TRUE)
FeaturePlot(data_seurat,reduction="umap",features=table_markers$gene[9:10],order=F,cols=c("#d9d9d933","#bd0026"), ncol = 2, label=TRUE)

FeaturePlot(data_seurat,reduction="umap",features=table_markers$gene[11:12],order=F,cols=c("#d9d9d933","#bd0026"), ncol = 2, label=TRUE)
FeaturePlot(data_seurat,reduction="umap",features=table_markers$gene[13:14],order=F,cols=c("#d9d9d933","#bd0026"), ncol = 2, label=TRUE)
FeaturePlot(data_seurat,reduction="umap",features=table_markers$gene[15:16],order=F,cols=c("#d9d9d933","#bd0026"), ncol = 2, label=TRUE)
FeaturePlot(data_seurat,reduction="umap",features=table_markers$gene[17:18],order=F,cols=c("#d9d9d933","#bd0026"), ncol = 2, label=TRUE)
FeaturePlot(data_seurat,reduction="umap",features=table_markers$gene[19:20],order=F,cols=c("#d9d9d933","#bd0026"), ncol = 2, label=TRUE)

FeaturePlot(data_seurat,reduction="umap",features=table_markers$gene[21:22],order=F,cols=c("#d9d9d933","#bd0026"), ncol = 2, label=TRUE)
FeaturePlot(data_seurat,reduction="umap",features=table_markers$gene[23:24],order=F,cols=c("#d9d9d933","#bd0026"), ncol = 2, label=TRUE)
FeaturePlot(data_seurat,reduction="umap",features=table_markers$gene[25:26],order=F,cols=c("#d9d9d933","#bd0026"), ncol = 2, label=TRUE)
FeaturePlot(data_seurat,reduction="umap",features=table_markers$gene[27:28],order=F,cols=c("#d9d9d933","#bd0026"), ncol = 2, label=TRUE)
FeaturePlot(data_seurat,reduction="umap",features=table_markers$gene[29:30],order=F,cols=c("#d9d9d933","#bd0026"), ncol = 2, label=TRUE)
```